Bayesian Hierarchical Modeling: An Introduction and Reassessment

### 2023-06-23 manuscript.Rmd
Updates based on Angelika's comments

@@

\Revision{There are several tutorials for readers interested in the frequentist approach of hierarchical model estimation} [e.g., @singmann2019introduction; @west2022linear]. \Revision{However, it has been argued that} the Bayesian framework offers a more intuitive approach for hierarchical modeling (Rouder & Lu, 2005; Lynch, 2007). The flexibility of the Bayesian approach facilitates the implementation of complex hierarchical models (Rouder, Morey, & Pratte, 2013). It also allows for the inclusion of prior knowledge into the analysis, as well as the monitoring of evidence for specific hypotheses as the data accumulate [@wagenmakers2018bayesian]. A complete discussion of the Bayesion approach is beyond the scope of this manuscript. We refer the interested reader to Etz and Vandekerckhove [-@etz2018introduction] and Wagenmakers and colleagues [-@wagenmakers2018bayesian].

@@

\begin{equation}
Y_{ijk} \sim \text{LogNormal}(\gamma_{i} + x_j \beta_{i} + u_j \delta_{7i} + v_j \delta_{6i} + w_j \delta_{4i} + z_j \delta_{3i}, \sigma^2).  
\end{equation}
Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution. \Revision{The likelihood function provides the likelihood of parameter values in the form of probability densities. The log-normal distribution only assigns probability to positive data values, and is, just as the RT distributions, right-skewed. Therefore, the log-normal distribution is a better match for the RT data.} The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model. However, it has a different interpretation. For example, for a location parameter of $\mu = 0$ and a scale parameter of $\sigma^2 = 1$, the mode of the log-normal distribution is $0.368$, and the mean is $1.65$.

@@

We follow the same procedure for prior specification in the log-normal model. As before, we prefer to apply conjugate priors. \Revision{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model} [@fink1997compendium]. \Revision{For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero. Applying this conjugate prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.} 
As a starting point, we can use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \@ref(fig:priorpredfigurepanel). These priors result in a slightly higher variance in RTs than we would expect, and, importantly, the mean response time differences are not centered around zero in all conditions. This pattern is inconsistent with our expectations, which means that prior distributions for the log-normal model need to be adjusted.

@@

The posterior distributions for the focal parameters in the normal model are presented in Figure \@ref(fig:modelposteriorsmufigure3). The posterior means estimated by `rstan` and `brms` can be found in [Online Supplement H](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/H%20-%20Posterior%20Distributions%20Tables). The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value [@wagenmakers2018bayesian].  \Revision{In contrast to the confidence interval, the credible interval provides the probability of a range of values for an effect. However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect} [@wagenmakers2020principle].

```{r geneftable, results = "hide", cache = TRUE, warning = FALSE}
# example - version 1 with SE 
rstanfithier8 <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
fitrstanhiertable <- summary(rstanfithier8, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2 <- as.data.frame(fitrstanhiertable)
rownamesfitrstanhiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
colnames(fitrstanhiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

@@

# example - version 2 without SE 
rstanfithier8 <- readRDS(file = "R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
fitrstanhiertable <- summary(rstanfithier8, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2 <- as.data.frame(fitrstanhiertable)
rownamesfitrstanhiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
colnames(fitrstanhiertable2) <- c("Mean", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

@@

For the digit effects, positive parameter estimates indicate that the response is slower for the non-baseline digits, that is, digits that are closer to 5. The parameters $\mu_{\delta_{6}}$ and $\mu_{\delta_{4}}$ are distributed around higher values. Note that this result is consistent with the symbolic distance hypothesis that the digits 4 and 6 have the highest response times because they are closest to the 5, the comparison value. The posterior distribution of the side effect $\mu_{\beta}$ is centered around zero with a posterior mean of `r printnum(fitrstanhiertable3["mu2", "Mean"], digits = 3)` and a 95% credible interval ranging from `r printnum(fitrstanhiertable3["mu2", "Lower Bound"], digits = 3)` to `r printnum(fitrstanhiertable3["mu2", "Upper Bound"], digits = 3)`. \Revision{Note that the credible interval overlaps with zero. Commonly, this is misinterpreted as an absence of an effect. However, in combination with the mean estimate, we can conclude that the effect is small.}

@@

\Revision{The individual variation can also be visualized in a different way with focusing on the relation between the individual estimates.} Figure \@ref(fig:modelestimatesfigure3) shows how the relationship between individuals' digit effects.^[The design and code for this figure are by @j_van_langen_2020_3715576.] The posterior means of individual effects are represented by the dots. The right side of each plot represents the distribution of the individual estimates. Note that this distribution is not the posterior distribution of the general effect. It rather corresponds to the left side of Figure 6.

@@

```{r geneftablelog, results = "hide", cache = TRUE, warning = F}
# example - with SE 
rstanfithier8log <- readRDS(file = "R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
fitrstanhiertablelog <- summary(rstanfithier8log, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2log <- as.data.frame(fitrstanhiertablelog)
rownamesfitrstanhiertablelog <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
colnames(fitrstanhiertable2log) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

@@

# example - without SE 
rstanfithier8log <- readRDS(file = "R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
fitrstanhiertablelog <- summary(rstanfithier8log, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2log <- as.data.frame(fitrstanhiertablelog)
rownamesfitrstanhiertablelog <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
colnames(fitrstanhiertable2log) <- c("Mean", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

@@

Same as the posterior distribution, Bayes factors can rarely be computed analytically due to the multidimensional integrals involved. The difficulty to compute Bayes factors is one of the major drawbacks of Bayesian inference. However, several algorithms are available to estimate Bayes factor from posterior samples. In the following section we will present two methods to obtain Bayes factors for the symbolic distance effect. \Revision{The first method that we will discuss, Savage-Dickey density ratio, evaluates the general effects with the individual effects in the model, while the second method, bridge-sampling, evaluates the full model, the general and individual effects are considered together. We want to make you aware that there have been discussions on which method should be preferred without consensus} [@van2021bayes; @van2023bayes; @rouder2023principles; @singmann2023statistics]. \Revision{Therefore, we will illustrate the use of the two approaches and discuss their potential.}

@@

\Revision{The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability. This is in line with our findings for the normal model. Finally, we can compare the normal model with the highest marginal likelihood to the log-normal with the highest marginal likelihood. The normal digit model from scenario 3 to the log-normal model from scenario 2. The log marginal likelihood for the normal digit model without individual differences was $6441.04$ while the log marginal likelihood for the log-normal digit model without individual differences was $10637.45$. This indicates that the Bayes factor is in favor of the log-normal digit model over the normal digit model.}

>>>>

\Revision{It is clear that neither of the alternative approaches yield satisfactory results by themselves. Conceptually, a hierarchical analysis provides a compromise between the aggregated and independent approach by drawing benefits from each method. The motivation for hierarchical models, as well as their general mathematical structure, has been covered by several tutorial articles} [e.g., @singmann2019introduction; @west2022linear]\Revision{, as well as text books (e.g., XX, XX).}

@@

The Bayesian framework offers a more intuitive approach for hierarchical modeling (Rouder & Lu, 2005; Lynch, 2007). The flexibility of the Bayesian approach facilitates the implementation of complex hierarchical models (Rouder, Morey, & Pratte, 2013). It also allows for the inclusion of prior knowledge into the analysis, as well as the monitoring of evidence for specific hypotheses as the data accumulate [@wagenmakers2018bayesian]. A complete discussion of the Bayesion approach is beyond the scope of this manuscript. We refer the interested reader to Etz and Vandekerckhove [-@etz2018introduction] and Wagenmakers and colleagues [-@wagenmakers2018bayesian].

@@

\begin{equation}
Y_{ijk} \sim \text{LogNormal}(\gamma_{i} + x_j \beta_{i} + u_j \delta_{7i} + v_j \delta_{6i} + w_j \delta_{4i} + z_j \delta_{3i}, \sigma^2).  
\end{equation}
Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution. \Revision{The likelihood function describes the probability model that is assumed to be underlying the data. When the parameters of this probability model are known, the joint probability of the data can be inferred from the likelihood function. The log-normal distribution only assigns probability to positive data values, and is, just as the RT distributions, right-skewed. Therefore, the log-normal distribution is a better match for the RT data.} The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model. However, it has a different interpretation. For example, for a location parameter of $\mu = 0$ and a scale parameter of $\sigma^2 = 1$, the mode of the log-normal distribution is $0.368$, and the mean is $1.65$.

@@

We follow the same procedure for prior specification in the log-normal model. As before, we prefer to apply conjugate priors. \Revision{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model} [@fink1997compendium]. \Revision{For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero. If the mean of the log-normal distribution is smaller than zero, this means that the peak of the log-normal distribution moves closer to zero, i.e., that response times are quicker but still larger than zero. When applying the normal distribution as prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.} 
As a starting point, we can use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \@ref(fig:priorpredfigurepanel). These priors result in a slightly higher variance in RTs than we would expect, and, importantly, the mean response time differences are not centered around zero in all conditions. This pattern is inconsistent with our expectations, which means that prior distributions for the log-normal model need to be adjusted.

@@

The posterior distributions for the focal parameters in the normal model are presented in Figure \@ref(fig:modelposteriorsmufigure3). The posterior means estimated by `rstan` and `brms` can be found in [Online Supplement H](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/H%20-%20Posterior%20Distributions%20Tables). The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value [@wagenmakers2018bayesian].  \Revision{In contrast to the confidence interval, the credible interval provides the probability that a parameter lies within a certain range. However, the credible interval does not provide evidence to conclude whether an effect is zero or not }[@wagenmakers2020principle].

```{r geneftable, results = "hide", cache = TRUE, warning = FALSE}
# example - version 1 with SE 
rstanfithier8 <- readRDS(file = "K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
fitrstanhiertable <- summary(rstanfithier8, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2 <- as.data.frame(fitrstanhiertable)
rownamesfitrstanhiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
colnames(fitrstanhiertable2) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

@@

# example - version 2 without SE 
rstanfithier8 <- readRDS(file = "K - R objects/rstan/normal model/hier_modelc_adj_07012022.rds")
fitrstanhiertable <- summary(rstanfithier8, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2 <- as.data.frame(fitrstanhiertable)
rownamesfitrstanhiertable <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
colnames(fitrstanhiertable2) <- c("Mean", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

@@

For the digit effects, positive parameter estimates indicate that the response is slower for the non-baseline digits, that is, digits that are closer to 5. The parameters $\mu_{\delta_{6}}$ and $\mu_{\delta_{4}}$ are distributed around higher values. Note that this result is consistent with the symbolic distance hypothesis that the digits 4 and 6 have the highest response times because they are closest to the 5, the comparison value. The posterior distribution of the side effect $\mu_{\beta}$ is centered around zero with a posterior mean of `r printnum(fitrstanhiertable3["mu2", "Mean"], digits = 3)` and a 95% credible interval ranging from `r printnum(fitrstanhiertable3["mu2", "Lower Bound"], digits = 3)` to `r printnum(fitrstanhiertable3["mu2", "Upper Bound"], digits = 3)`. \Revision{Note that the credible interval overlaps with zero. Commonly, this is misinterpreted as an absence of an effect. However, credible intervals should only be used for parameter estimation. In cases where the credible overlaps with zero,  the effect is likely small, but can be non-zero.}

@@

\Revision{The individual variation can also be visualized differently, for example by connecting individual estimates for each parameter. This type of variation can reveal correlations between individual parameter estimates, for example, participants scoring high on one digit parameter may also score high on another digit parameter.} Figure \@ref(fig:modelestimatesfigure3) shows the relationship between individuals' digit effects.^[The design and code for this figure are by @j_van_langen_2020_3715576.] The posterior means of individual effects are represented by the dots. The right side of each plot represents the distribution of the individual estimates. Note that this distribution is not the posterior distribution of the general effect. It rather corresponds to the left side of Figure 6.

@@

```{r geneftablelog, results = "hide", cache = TRUE, warning = F}
# example - with SE 
rstanfithier8log <- readRDS(file = "K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
fitrstanhiertablelog <- summary(rstanfithier8log, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "se_mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2log <- as.data.frame(fitrstanhiertablelog)
rownamesfitrstanhiertablelog <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
colnames(fitrstanhiertable2log) <- c("Mean", "SE", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

@@

# example - without SE 
rstanfithier8log <- readRDS(file = "K - R objects/rstan/log-normal model/hier_logmodelc_adj13012022.rds")
fitrstanhiertablelog <- summary(rstanfithier8log, pars = c("mu", "mu2", "mu3", "mu4", "mu5", "mu6"))$summary[,c("mean", "2.5%", "97.5%", "n_eff", "Rhat")]
# as.data.frame(fitrstanhiertable)
fitrstanhiertable2log <- as.data.frame(fitrstanhiertablelog)
rownamesfitrstanhiertablelog <-  c("$\\mu_{\\gamma}$","$\\mu_{\\beta}$", "$\\mu_{\\delta_{7}}$", "$\\mu_{\\delta_{6}}$", "$\\mu_{\\delta_{4}}$", "$\\mu_{\\delta_{3}}$")
colnames(fitrstanhiertable2log) <- c("Mean", "Lower Bound", "Upper Bound", "$n_{eff}$","$\\hat{R}$")

@@

Same as the posterior distribution, Bayes factors can rarely be computed analytically due to the multidimensional integrals involved. The difficulty to compute Bayes factors is one of the major drawbacks of Bayesian inference. However, several algorithms are available to estimate Bayes factor from posterior samples. In the following section we will present two methods to obtain Bayes factors for the symbolic distance effect. \Revision{The first method that we discuss, Savage-Dickey density ratio, tests the general effect while disregarding individual variation (similar to type III sums of squares in ANOVA models). This means that the compared models differ in only a single parameter, that is, the general effect parameter, and individual effects are present in both models. The Savage-Dickey density ratio is computationally efficient, but can only be used for tests of a single parameter value as described before. This is why we additionally introduce a second approach, bridge-sampling, that is less computationally efficient, but more versatile. In this manuscript, we will use bridge sampling to test the general effects and individual variation together, as well as to compare the predictive accuracy of the normal and log-normal model. There have been discussions on which method should be preferred without consensus} [@van2021bayes; @van2023bayes; @rouder2023principles; @singmann2023statistics]. \Revision{Therefore, we illustrate the use of the two approaches and discuss their potential.}

@@

\Revision{The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability. This is in line with our findings for the normal model. Finally, we can compare the normal model with the highest marginal likelihood to the log-normal with the highest marginal likelihood. The normal digit model from scenario 3 to the log-normal model from scenario 2. The log marginal likelihood for the normal digit model without individual differences was $6441.04$ while the log marginal likelihood for the log-normal digit model without individual differences was $10637.45$. This indicates that the Bayes factor is strongly in favor of the log-normal digit model over the normal digit model.}

### 2023-06-23 manuscript.tex
Updates based on Angelika's comments

@@

\textcolor{blue}{There are several tutorials for readers interested in the frequentist approach of hierarchical model estimation} (e.g., Singmann \& Kellen, 2019; West, Welch, \& Galecki, 2022). \textcolor{blue}{However, it has been argued that} the Bayesian framework offers a more intuitive approach for hierarchical modeling (Rouder \& Lu, 2005; Lynch, 2007). The flexibility of the Bayesian approach facilitates the implementation of complex hierarchical models (Rouder, Morey, \& Pratte, 2013). It also allows for the inclusion of prior knowledge into the analysis, as well as the monitoring of evidence for specific hypotheses as the data accumulate (Wagenmakers et al., 2018). A complete discussion of the Bayesion approach is beyond the scope of this manuscript. We refer the interested reader to Etz and Vandekerckhove (2018) and Wagenmakers and colleagues (2018).

@@

\begin{equation}
Y_{ijk} \sim \text{LogNormal}(\gamma_{i} + x_j \beta_{i} + u_j \delta_{7i} + v_j \delta_{6i} + w_j \delta_{4i} + z_j \delta_{3i}, \sigma^2).  
\end{equation}
Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution. \textcolor{blue}{The likelihood function provides the likelihood of parameter values in the form of probability densities. The log-normal distribution only assigns probability to positive data values, and is, just as the RT distributions, right-skewed. Therefore, the log-normal distribution is a better match for the RT data.} The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model. However, it has a different interpretation. For example, for a location parameter of \(\mu = 0\) and a scale parameter of \(\sigma^2 = 1\), the mode of the log-normal distribution is \(0.368\), and the mean is \(1.65\).

@@

We follow the same procedure for prior specification in the log-normal model. As before, we prefer to apply conjugate priors. \textcolor{blue}{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model} (Fink, 1997b). \textcolor{blue}{For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero. Applying this conjugate prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.}
As a starting point, we can use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \ref{fig:priorpredfigurepanel}. These priors result in a slightly higher variance in RTs than we would expect, and, importantly, the mean response time differences are not centered around zero in all conditions. This pattern is inconsistent with our expectations, which means that prior distributions for the log-normal model need to be adjusted.

@@

The posterior distributions for the focal parameters in the normal model are presented in Figure \ref{fig:modelposteriorsmufigure3}. The posterior means estimated by \texttt{rstan} and \texttt{brms} can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/H\%20-\%20Posterior\%20Distributions\%20Tables}{Online Supplement H}. The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value (Wagenmakers et al., 2018). \textcolor{blue}{In contrast to the confidence interval, the credible interval provides the probability of a range of values for an effect. However, considering the prior and under the assumption that there is an effect, the credible interval does not provide evidence to conclude whether there is an effect} (Wagenmakers, Lee, Rouder, \& Morey, 2020).

@@

For the digit effects, positive parameter estimates indicate that the response is slower for the non-baseline digits, that is, digits that are closer to 5. The parameters \(\mu_{\delta_{6}}\) and \(\mu_{\delta_{4}}\) are distributed around higher values. Note that this result is consistent with the symbolic distance hypothesis that the digits 4 and 6 have the highest response times because they are closest to the 5, the comparison value. The posterior distribution of the side effect \(\mu_{\beta}\) is centered around zero with a posterior mean of -0.007 and a 95\% credible interval ranging from -0.046 to 0.031. \textcolor{blue}{Note that the credible interval overlaps with zero. Commonly, this is misinterpreted as an absence of an effect. However, in combination with the mean estimate, we can conclude that the effect is small.}

@@

\textcolor{blue}{The individual variation can also be visualized in a different way with focusing on the relation between the individual estimates.} Figure \ref{fig:modelestimatesfigure3} shows how the relationship between individuals' digit effects.\footnote{The design and code for this figure are by Langen (2020).} The posterior means of individual effects are represented by the dots. The right side of each plot represents the distribution of the individual estimates. Note that this distribution is not the posterior distribution of the general effect. It rather corresponds to the left side of Figure 6.

@@

Same as the posterior distribution, Bayes factors can rarely be computed analytically due to the multidimensional integrals involved. The difficulty to compute Bayes factors is one of the major drawbacks of Bayesian inference. However, several algorithms are available to estimate Bayes factor from posterior samples. In the following section we will present two methods to obtain Bayes factors for the symbolic distance effect. \textcolor{blue}{The first method that we will discuss, Savage-Dickey density ratio, evaluates the general effects with the individual effects in the model, while the second method, bridge-sampling, evaluates the full model, the general and individual effects are considered together. We want to make you aware that there have been discussions on which method should be preferred without consensus} (Doorn, Aust, Haaf, Stefan, \& Wagenmakers, 2021, 2023; Rouder, Schnuerch, Haaf, \& Morey, 2023; Singmann et al., 2023). \textcolor{blue}{Therefore, we will illustrate the use of the two approaches and discuss their potential.}

@@

\textcolor{blue}{The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability. This is in line with our findings for the normal model. Finally, we can compare the normal model with the highest marginal likelihood to the log-normal with the highest marginal likelihood. The normal digit model from scenario 3 to the log-normal model from scenario 2. The log marginal likelihood for the normal digit model without individual differences was $6441.04$ while the log marginal likelihood for the log-normal digit model without individual differences was $10637.45$. This indicates that the Bayes factor is in favor of the log-normal digit model over the normal digit model.}

>>>>

\textcolor{blue}{It is clear that neither of the alternative approaches yield satisfactory results by themselves. Conceptually, a hierarchical analysis provides a compromise between the aggregated and independent approach by drawing benefits from each method. The motivation for hierarchical models, as well as their general mathematical structure, has been covered by several tutorial articles} (e.g., Singmann \& Kellen, 2019; West, Welch, \& Galecki, 2022)\textcolor{blue}{, as well as text books (e.g., XX, XX).}

@@

The Bayesian framework offers a more intuitive approach for hierarchical modeling (Rouder \& Lu, 2005; Lynch, 2007). The flexibility of the Bayesian approach facilitates the implementation of complex hierarchical models (Rouder, Morey, \& Pratte, 2013). It also allows for the inclusion of prior knowledge into the analysis, as well as the monitoring of evidence for specific hypotheses as the data accumulate (Wagenmakers et al., 2018). A complete discussion of the Bayesion approach is beyond the scope of this manuscript. We refer the interested reader to Etz and Vandekerckhove (2018) and Wagenmakers and colleagues (2018).

@@

\begin{equation}
Y_{ijk} \sim \text{LogNormal}(\gamma_{i} + x_j \beta_{i} + u_j \delta_{7i} + v_j \delta_{6i} + w_j \delta_{4i} + z_j \delta_{3i}, \sigma^2).  
\end{equation}
Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution. \textcolor{blue}{The likelihood function describes the probability model that is assumed to be underlying the data. When the parameters of this probability model are known, the joint probability of the data can be inferred from the likelihood function. The log-normal distribution only assigns probability to positive data values, and is, just as the RT distributions, right-skewed. Therefore, the log-normal distribution is a better match for the RT data.} The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model. However, it has a different interpretation. For example, for a location parameter of \(\mu = 0\) and a scale parameter of \(\sigma^2 = 1\), the mode of the log-normal distribution is \(0.368\), and the mean is \(1.65\).

@@

We follow the same procedure for prior specification in the log-normal model. As before, we prefer to apply conjugate priors. \textcolor{blue}{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model} (Fink, 1997b). \textcolor{blue}{For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero. If the mean of the log-normal distribution is smaller than zero, this means that the peak of the log-normal distribution moves closer to zero, i.e., that response times are quicker but still larger than zero. When applying the normal distribution as prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.}
As a starting point, we can use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \ref{fig:priorpredfigurepanel}. These priors result in a slightly higher variance in RTs than we would expect, and, importantly, the mean response time differences are not centered around zero in all conditions. This pattern is inconsistent with our expectations, which means that prior distributions for the log-normal model need to be adjusted.

@@

The posterior distributions for the focal parameters in the normal model are presented in Figure \ref{fig:modelposteriorsmufigure3}. The posterior means estimated by \texttt{rstan} and \texttt{brms} can be found in \href{https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/tree/main/H\%20-\%20Posterior\%20Distributions\%20Tables}{Online Supplement H}. The credible interval displayed in the figure is the Bayesian version of a confidence interval and provides a measure of uncertainty about the parameter value (Wagenmakers et al., 2018). \textcolor{blue}{In contrast to the confidence interval, the credible interval provides the probability that a parameter lies within a certain range. However, the credible interval does not provide evidence to conclude whether an effect is zero or not }(Wagenmakers, Lee, Rouder, \& Morey, 2020).

@@

For the digit effects, positive parameter estimates indicate that the response is slower for the non-baseline digits, that is, digits that are closer to 5. The parameters \(\mu_{\delta_{6}}\) and \(\mu_{\delta_{4}}\) are distributed around higher values. Note that this result is consistent with the symbolic distance hypothesis that the digits 4 and 6 have the highest response times because they are closest to the 5, the comparison value. The posterior distribution of the side effect \(\mu_{\beta}\) is centered around zero with a posterior mean of -0.007 and a 95\% credible interval ranging from -0.046 to 0.031. \textcolor{blue}{Note that the credible interval overlaps with zero. Commonly, this is misinterpreted as an absence of an effect. However, credible intervals should only be used for parameter estimation. In cases where the credible overlaps with zero,  the effect is likely small, but can be non-zero.}

@@

\textcolor{blue}{The individual variation can also be visualized differently, for example by connecting individual estimates for each parameter. This type of variation can reveal correlations between individual parameter estimates, for example, participants scoring high on one digit parameter may also score high on another digit parameter.} Figure \ref{fig:modelestimatesfigure3} shows the relationship between individuals' digit effects.\footnote{The design and code for this figure are by Langen (2020).} The posterior means of individual effects are represented by the dots. The right side of each plot represents the distribution of the individual estimates. Note that this distribution is not the posterior distribution of the general effect. It rather corresponds to the left side of Figure 6.

@@

Same as the posterior distribution, Bayes factors can rarely be computed analytically due to the multidimensional integrals involved. The difficulty to compute Bayes factors is one of the major drawbacks of Bayesian inference. However, several algorithms are available to estimate Bayes factor from posterior samples. In the following section we will present two methods to obtain Bayes factors for the symbolic distance effect. \textcolor{blue}{The first method that we discuss, Savage-Dickey density ratio, tests the general effect while disregarding individual variation (similar to type III sums of squares in ANOVA models). This means that the compared models differ in only a single parameter, that is, the general effect parameter, and individual effects are present in both models. The Savage-Dickey density ratio is computationally efficient, but can only be used for tests of a single parameter value as described before. This is why we additionally introduce a second approach, bridge-sampling, that is less computationally efficient, but more versatile. In this manuscript, we will use bridge sampling to test the general effects and individual variation together, as well as to compare the predictive accuracy of the normal and log-normal model. There have been discussions on which method should be preferred without consensus} (Doorn, Aust, Haaf, Stefan, \& Wagenmakers, 2021, 2023; Rouder, Schnuerch, Haaf, \& Morey, 2023; Singmann et al., 2023). \textcolor{blue}{Therefore, we illustrate the use of the two approaches and discuss their potential.}

@@

\textcolor{blue}{The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability. This is in line with our findings for the normal model. Finally, we can compare the normal model with the highest marginal likelihood to the log-normal with the highest marginal likelihood. The normal digit model from scenario 3 to the log-normal model from scenario 2. The log marginal likelihood for the normal digit model without individual differences was $6441.04$ while the log marginal likelihood for the log-normal digit model without individual differences was $10637.45$. This indicates that the Bayes factor is strongly in favor of the log-normal digit model over the normal digit model.}
